Map/Reduce for the common man

### 2010-02-16 doc/slides.tex
*** empty log message ***

@@

\begin{frame}[fragile]\frametitle{API}
%\footnotesize
  \begin{itemize}
  \item traditional \texttt{List.map}
\begin{verbatim}
  val map : 
    f:('a -> 'b) -> 'a list -> 'b list
\end{verbatim}

@@

\begin{verbatim}
  val map_reduce :
    map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
    'c -> 'a list -> 'c
\end{verbatim}
    \begin{displaymath}
      \mathtt{map\_reduce}~f~g~c~l
      \begin{array}[t]{l}
        = \mathtt{fold}~g~c~(\mathtt{map}~f~l) \\
        = \mathtt{fold}~(\mathtt{fun}~c~x\rightarrow g~c~(f~x))~c~l\\
      \end{array}
    \end{displaymath}
  \end{itemize}
\end{frame}

\begin{frame}[fragile]\frametitle{}
\begin{verbatim}
val map_reduce_ac :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b

val map_reduce_a :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
\end{verbatim}
\end{frame}

\begin{frame}\frametitle{Options}
  \begin{center}
    \begin{tabular}{|c|c|c|c|c|c|}
      \hline
                 & \multicolumn{2}{|c|}{map} &
                 \multicolumn{3}{|c|}{map/reduce} \\\cline{2-6}
                 & map & local & remote & ac & a \\\hline\hline
      sequential & & & & & \\\hline
      cores      & & & & & \\\hline
      network    & & & & & \\\hline
    \end{tabular}
  \end{center}
\end{frame}

\begin{frame}\frametitle{}
  first demo
\end{frame}

\begin{frame}\frametitle{}
  detailed API
\end{frame}

\begin{frame}\frametitle{}
  another, more impressive demo ?
\end{frame}

>>>>

\begin{frame}\frametitle{}
  TODO: introduce cores/network idea
\end{frame}

@@

\begin{frame}\frametitle{API}
%\footnotesize
  \begin{itemize}
  \item traditional \texttt{List.map}
\begin{ocaml}
  val map : 
    f:('a -> 'b) -> 'a list -> 'b list
\end{ocaml}

@@

\begin{ocaml}
  val map_reduce :
    map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
    'c -> 'a list -> 'c
\end{ocaml}
    \begin{displaymath}
      \mathtt{map\_reduce}~f~g~c~l
      \begin{array}[t]{l}
        = \mathtt{fold}~g~c~(\mathtt{map}~f~l) \\
        = \mathtt{fold}~(\mathtt{fun}~c~x\rightarrow g~c~(f~x))~c~l\\
      \end{array}
    \end{displaymath}
    assuming $g$ is not sensitive to element ordering
  \end{itemize}

  example :
  \begin{displaymath}
    \sum_{x\in l}f(x) = \mathtt{map\_reduce}~f~(+)~0~l
  \end{displaymath}
\end{frame}

\begin{frame}\frametitle{Special Case}
  \emph{idea}: reduce as early as possible, as in
  \begin{displaymath}
    \sum_{x\in l}f(x) = \mathtt{map\_reduce}~f~(+)~0~l
  \end{displaymath}
  in this case, + is \emph{associative / commutative}, so reductions are
  order-insensitive and can be performed as soon as two intermediate
  results are available

  \begin{ocaml}
val map_reduce_ac :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
  \end{ocaml}
\end{frame}

\begin{frame}\frametitle{Special Case 2}

  \emph{order-sensitive} computation:
  \begin{displaymath}
    \mathtt{map\_reduce}~\mathtt{String.uppercase}~(\texttt{\^{}})~\texttt{""}~
    \texttt{["h"; "e"; "l"; "l"; "o"]}
  \end{displaymath}

  operation is only \emph{associative}, yet we can reduce as soon as
  adjacent intermediate results are available

\begin{ocaml}
val map_reduce_a :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
\end{ocaml}
\end{frame}

\begin{frame}\frametitle{Local/Remote reductions}
  reductions can be performed either by the master (local) or by the
  workers (remote)

  \bigskip
  remote reductions allow other computations to be performed in parallel

  \texttt{map\_reduce} takes the following forms
  \begin{ocaml}
val map_local_reduce :
  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
  'c -> 'a list -> 'c
val map_remote_reduce :
  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
  'c -> 'a list -> 'c
  \end{ocaml}
\end{frame}

\begin{frame}\frametitle{API Summary}
  \begin{ocaml}
val map : 
  f:('a -> 'b) -> 'a list -> 'b list
val map_local_reduce :
  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
  'c -> 'a list -> 'c
val map_remote_reduce :
  map:('a -> 'b) -> reduce:('c -> 'b -> 'c) -> 
  'c -> 'a list -> 'c
val map_reduce_ac :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
val map_reduce_a :
  map:('a -> 'b) -> reduce:('b -> 'b -> 'b) -> 
  'b -> 'a list -> 'b
  \end{ocaml}
\end{frame}

\begin{frame}\frametitle{Library Options}

  3 modules, with 5 functions each

  \begin{center}
    \begin{tabular}{|l|c|c|c|c|c|}
      \hline
                 & \multicolumn{2}{|c|}{map} &
                 \multicolumn{3}{|c|}{map/reduce} \\\cline{2-6}
                 & map & local & remote & ~ac~ & ~~a~~ \\\hline\hline
      \texttt{Sequential} & & & \multicolumn{3}{|c|}{(no difference)} \\\hline
      \texttt{Cores}      & & & & & \\\hline
      \texttt{Network}    & & & & & \\\hline
    \end{tabular}
  \end{center}
  % TODO: fill the cells
\end{frame}

\begin{frame}\frametitle{Demo}
  \emph{demo}: solving the $n$-queens puzzle
\end{frame}

\begin{frame}\frametitle{Demo}
  \emph{demo}: solving the $n$-queens puzzle

  example: $n=16$
  \begin{center}
    \begin{tabular}{|l|r|}
      \hline
      sequential & 24.4 s \\\hline
      2 cores    & 12.6 s \\\hline
      12 cores on moloch & 3.0 s \\\hline
    \end{tabular}
  \end{center}

  example: $n=17$
  \begin{center}
    \begin{tabular}{|l|r|}
       \hline
      sequential & 172.0 s \\\hline
      moloch (12), orcus (4), localhost(1) & 11.7 s \\\hline
    \end{tabular}
  \end{center}
\end{frame}
